What We Eval in the Shadows

### 2021-08-09 paper/main.tex
Remove base and kaggle from Section 4

\mypara{CRAN.} Packages hosted on the Comprehensive R Archive Network ({\small
  \url{cran.r-project.org}}, aka CRAN) have three sources of runnable: unit
\emph{tests} for individual functions, code \emph{examples} embedded in the
documentation, and use-cases called \emph{vignettes}. All runnable snippets can
be extracted into independent files. We download \CranPackages packages and
extract \CranRunnableScripts scripts with \CranRunnableCode lines of code
(\CranRunnableCodeExamplesRnd in examples, \CranRunnableCodeVignettesRnd in
vignettes and \CranRunnableCodeTestsRnd in tests). CRAN packages are authored by
experienced developers.

@@

       \node [block, fill = red, draw = darkred, very thick] (download) {
         \nodename{
           \vspace{1mm}\textbf{Download}\vspace{1mm}\\
           15K Packages\\
           8K Kaggle Kernels\\
           219 Kaggle Datasets
         }
       };
       \node [below = \textsep of download]   (downloaddesc)   {\nodedesc{12 Hours}};
       \node [block, right = \nodesep of download, fill = yellow, draw = darkyellow, very thick] (extract) {
         \nodename{
           \vspace{1mm}\textbf{Extract}\vspace{1mm}\\
           240K package programs\\%, 4.6M LOC\\
           74 Kaggle kernels\\%, 13K LOC\\
           \newline
         }
       };
       \node [below = \textsep of extract]   (extractdesc)   {\nodedesc{2 Hours}};
       \node [block, right = \nodesep of extract, fill = blue, draw = darkblue, very thick] (trace) {
         \nodename{
           \vspace{1mm}\textbf{Trace}\vspace{1mm}\\
           49M eval calls \\
           14M side-effects\\
           \newline
         }
       };
       \node [below = \textsep of trace]     (tracedesc)     {\nodedesc{1 Day 12 Hours}};
       \node [block, right = \nodesep of trace, fill = green, draw = darkgreen, very thick] (analyze) {
         \nodename{
           \vspace{1mm}\textbf{Analyze}\vspace{1mm}\\
           2.8M unique eval calls\\
           22\% packages use eval\\
           \newline
         }
       };
       \node [below = \textsep of analyze]    (analyzedesc)    {\nodedesc{2 Hours}};
       \draw [connector] (download)   edge (extract);
       \draw [connector] (extract)    edge (trace);
       \draw [connector] (trace)      edge (analyze);
     \end{tikzpicture}
   }
   \vspace{-2mm}\caption{Pipeline}\label{fig:pipeline}
 \end{figure}

@@

\noindent The pipeline runs in parallel on three servers, each with 2.3GHz Intel
Xeon 6140 processor with 72 cores and 256GB of RAM, and shared OCFS network
storage. Servers run identical docker images with all dependencies installed.

@@

\paragraph{CRAN}

@@

\paragraph{Base}\label{sec:usage-base}
We recorded \baseAllcalls \eval calls in the base libraries. This
comes from running 10\% randomly selected programs from the
\CranRunnableScriptsRnd extracted programs from CRAN packages, covering
\baseTriggeredevalpct of the \BaseEvalCallSites sites. Most of the calls
(\baseEvalsratio) are to the \c{eval} function, and most arguments
(\baseCodepercent) are expressions. However, unlike CRAN, the majority
(\baseLanguagepercent) of the arguments are language objects. The median
argument size is \baseMedianszeval, which is more than for CRAN and the maximum
size is \baseMaxszeval, less than for CRAN. Small instruction counts ($<50$)
amount to \baseSmalleventspct of calls. A single site in the \c{match.arg}
function is responsible for \baseTopFuncPercent of all recorded calls. This
function provides a convenient way for argument verification using partial
matching. It is heavily relied on in R functions that use string arguments to
parameterize function behavior, common practice in R. For example, in a body of
\c{center<-function(x,type=c("mean","median","trimmed"))} one can get the
value of \c{type} parameter using \c{type<-match.arg(type)}. A call
\c{center(x,"trim")} will match \c{type} to \c{"trimmed"}. The challenge
dealing with Base is that every program uses it. It can thus generate extreme
amounts of data, but the data is quite predictable. From now on, we only mention
Base when it surprises.

\paragraph{Kaggle}

In total, \kaggleAllcalls \eval were recorded, all to \eval function. Out of
\kaggleStaticeval sites from \KaggleWithEvals scripts, only \kaggleTriggeredeval
sites in \kaggleNbruns scripts were hit; \KaggleFailedScripts scripts failed to
run, and the others did not exercise the \eval sites. This is partially expected
as Kaggle code does not need to abide by any checks. Upon a manual inspection,
we observed that indeed, the failing scripts were of poor quality, often not
finished, using misspelled package names, hard-coded file paths, or accessing
missing files. Most call sites are invoked once. Only one site is called
\kaggleMaxcalls times. Expressions amount for \kaggleCodepercent of arguments,
there are very few symbols (\kaggleSymbolpercent). Most expressions result from
calls to \c{parse}; thus, most \evals start with strings. The median argument
size is \kaggleMedianszeval, which makes sense, as few arguments are only
symbols. The largest argument is \kaggleMaxszeval. The distribution of
instructions per eval is similar to CRAN; \kaggleSmalleventspct of \evals
execute fewer than 50 instructions. Manual inspection of \eval usage in Kaggle
suggests that it is consistent with the data obtained for CRAN. We do not
discuss it further.

\paragraph{Discussion}
\Eval is central to R's implementation as it is omnipresent in the Base library.
That library is, in turn, used by every single package. To avoid Base dominating
the result, we excluded calls to it from the results. CRAN Packages, which are
typically developed by experienced programmers, make regular and varied use of
\eval. They represent our most interesting data set as these packages result
from over 20 years of contributions by thousands of authors. The code in
packages is well maintained and relatively well tested. Finally, Kaggle scripts
are often written by less sophisticated users and likely perform more
straightforward tasks, thus having a lesser need for \eval, and its uses
originate from strings as these are easier to manipulate for end users.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Taxonomy of Eval}

>>>>

\mypara{CRAN.} Packages hosted on the Comprehensive R Archive Network ({\small
\url{cran.r-project.org}}, aka CRAN) have three sources of runnable: unit
\emph{tests} for individual functions, code \emph{examples} embedded in the
documentation, and use-cases called \emph{vignettes}. All runnable snippets
can be extracted into independent files. We download \CranPackages packages and
extract \CranRunnableScripts scripts with \CranRunnableCode lines of code
(\CranRunnableCodeExamplesRnd in examples, \CranRunnableCodeVignettesRnd in
vignettes and \CranRunnableCodeTestsRnd in tests). CRAN packages contain 20
years of contribution from thousands from thousands of authors, often
experienced R developers.

@@

       \node [block, fill = red, draw = darkred, very thick] (download) {
         \nodename{
           \vspace{1mm}\textbf{Download}\vspace{1mm}\\
           15K packages\\
           8K Kaggle kernels\\
           219 Kaggle datasets
         }
       };
       \node [below = \textsep of download]   (downloaddesc)   {\nodedesc{12 hours}};
       \node [block, right = \nodesep of download, fill = yellow, draw = darkyellow, very thick] (extract) {
         \nodename{
           \vspace{1mm}\textbf{Extract}\vspace{1mm}\\
           240K package programs\\%, 4.6M LOC\\
           74 Kaggle kernels\\%, 13K LOC\\
           \newline
         }
       };
       \node [below = \textsep of extract]   (extractdesc)   {\nodedesc{2 hours}};
       \node [block, right = \nodesep of extract, fill = blue, draw = darkblue, very thick] (trace) {
         \nodename{
           \vspace{1mm}\textbf{Trace}\vspace{1mm}\\
           49M eval calls \\
           14M side-effects\\
           \newline
         }
       };
       \node [below = \textsep of trace]     (tracedesc)     {\nodedesc{36 hours}};
       \node [block, right = \nodesep of trace, fill = green, draw = darkgreen, very thick] (analyze) {
         \nodename{
           \vspace{1mm}\textbf{Analyze}\vspace{1mm}\\
           2.8M unique eval calls\\
           22\% packages use eval\\
           \newline
         }
       };
       \node [below = \textsep of analyze]    (analyzedesc)    {\nodedesc{2 hours}};
       \draw [connector] (download)   edge (extract);
       \draw [connector] (extract)    edge (trace);
       \draw [connector] (trace)      edge (analyze);
     \end{tikzpicture}
   }
   \vspace{-2mm}\caption{Pipeline}\label{fig:pipeline}
 \end{figure}

@@

\vspace{1mm} 
\noindent The pipeline runs in parallel~\cite{GNUparallel}
orchestrated by a Makefile. Servers have identical environments thanks to a
docker image with all dependencies installed\footnote{The docker image is part
  of the accompanied artifact available at\\
\url{https://github.com/PRL-PRG/evalr-experiment/tree/oopsla21}}.

@@

@@

\paragraph{Summary}

\Eval is used in CRAN packages. Nearly every fourth package contains at least
one \eval site. However, most packages that use \eval do that modestly. Half of
them contain less than three \eval sites. Packages that rely heavily on \eval
are mostly the ones that provide functions for statistical modeling and
simulation. At runtime, over third of extracted programs triggered calls to
\eval. Half of the \packageAllcallsRnd recorded calls come from a single \eval
site, but for the rest, most of the packages have low \eval call frequency.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Taxonomy of \Eval}
